The error when inferring phylogenies with incipient species by a Birth-Death model

### 2018-04-06 article.tex
Data set specified

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Abbreviations}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{l l}
    \hline
    Symbol & Description \\
    \hline
    \hline
    $b_g$ & Speciation initiation rate of a good species (probability per lineage, per million year) \\
    $b_i$ & Speciation initiation rate of an incipient species (probability per lineage, per million year) \\
    $\lambda$ & Speciation completion rate (probability per lineage, per million year) \\
    $\mu_g$ & Extinction rate of a good species (probability per lineage, per million year) \\
    $\mu_i$ & Extinction rate of an incipient species (probability per lineage, per million year) \\
    \hline
    $t_c$ & Crown age (million years) \\
    $r$ & Mutation rate (probability of mutation per nucleotide per million years) \\
    $l_a$ & DNA alignment length (base pairs) \\
    \hline
    $n_i$ & Number of incipient species trees per parameter set \\
    $n_a$ & Number of alignments per species tree \\
    $n_b$ & Number of BEAST2 runs per alignment \\
    \hline
    $n_s$ & Number of MCMC samples \\
    $i_s$ & MCMC sampling interval \\
    \hline
    $n_{ti}$ & Number of taxa in an incipient species tree \\
    $n_{t}$ & Number of taxa in a (good) species tree \\
    $t_{\mean{ds}}$ & Average duration of speciation \\
    $t_{sc}$ & Speciation completion time \\
    $\Delta_{nLTT}$ & nLTT statistic value \\
    $\mean{\Delta_{nLTT}}$ & Average nLTT statistic value of a posterior \\
    $\median{\Delta_{nLTT}}$ & Median nLTT statistic value of a posterior \\
    \hline
  \end{tabular}
  \caption{
    Parameter descriptions. The sections are (1) the biological parameters, (2) the
    biological tuning parameters, (3) parameters involved in replication, (4) MCMC
    tuning parameters, and (5) other symbols.
  }
  \label{table:parameter_descriptions}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Parameters}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{l l}
    \hline
    Parameter             & Values \\
    \hline
    \hline
    $b = b_g = b_i$       & 0.1, 0.5, 1.0 \\
    $\lambda$             & 0.1, 0.3, 1.0, $10^6$ \\
    $\mu = \mu_g = \mu_i$ & 0.0, 0.1, 0.2, 0.4 \\
    \hline
    $t_c$                 & 15 \\
    $r$                   & 1000 \\
    $l_a$                 & $15000$ \\
    \hline
    $n_i$                 & 178 \\
    $n_a$                 & 2 per species tree\\
    $n_b$                 & 2 per alignment\\
    \hline
    $n_s$                 & $10^3$ states (sampled)\\
    $i_s$                 & $10^3$ states (between samples)\\
    \hline
  \end{tabular}
  \caption{
    Parameter values. The sections are (1) the biological parameters, (2) the
    biological tuning parameters, (3) parameters involved in replication, 
    and (4) MCMC tuning parameters.
  }
  \label{table:parameter_values}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data set specifications}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The raw data file will consist of 40K rows. On each row there will be 1012 values:

\begin{itemize}  
  \item 12 simulation parameters:
  \begin{enumerate}  
    \item speciation initiation rate
    \item speciation completion rate
    \item extinction rate
    \item crown age
    \item crown age sigma
    \item sampling method
    \item mutation rate
    \item sequence length
    \item mcmc length
    \item tree sim rng seed
    \item alignment rng seed
    \item beast2 rnd seed
  \end{enumerate}
  \item 1000 nLTT statistics
\end{itemize}

The summarized data file will consist of 40K rows. On each row there will be 29 values:

\begin{itemize}  
  \item 12 simulation parameters:
  \begin{enumerate}  
    \item speciation initiation rate
    \item speciation completion rate
    \item extinction rate
    \item crown age
    \item crown age sigma
    \item sampling method
    \item mutation rate
    \item sequence length
    \item mcmc length
    \item tree sim rng seed
    \item alignment rng seed
    \item beast2 rnd seed
  \end{enumerate}
  \item [NOTE: add yes/no?] expected speciation completion time (from Etienne \& Rosindell 2012)
  \item [NOTE: add yes/no?] expected duration of speciation (from Etienne \& Rosindell 2012)
  \item number of taxa in incipient species tree (also: number of subspecies)
  \item [NOTE: add yes/no?] would there be a difference in branch lengths between the two sampling methods?
  \item number of taxa in species tree (also: number of species)
  \item mean nLTT statistic
  \item median nLTT statistic
  \item lower bound of 95\% HPD value of nLTT statistic
  \item upper bound of 95\% HPD value of nLTT statistic
  \item 11 ESSes of all parameters estimated by BEAST2:
  \begin{enumerate}  
    \item posterior
    \item likelihood 
    \item prior 
    \item treeLikelihood 
    \item TreeHeight 
    \item BirthDeath 
    \item BDBirthRate 
    \item BDDeathRate 
    \item logP.mrca 
    \item mrcatime 
    \item clockRate
  \end{enumerate}
  \item [some] MCMC operator acceptances
\end{itemize}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{l l l}
    \hline
    Parameter             & Description & Values \\
    \hline
    \hline
    $b = b_g = b_i$       & Speciation initiation rate & 0.1, 0.5, 1.0 \\
    $\lambda$             & Speciation completion rate & 0.1, 0.3, 1.0, $10^6$ \\
    $\mu = \mu_g = \mu_i$ & Extinction rate & 0.0, 0.1, 0.2, 0.4 \\
    $t_c$                 & Crown age & 15 \\
    $\sigma_c$            & Standard deviation around crown age & 0.001 \\
    M                     & Sampling method & Youngest, oldest \\
    $r$                   & Mutation rate & 1000 \\
    $l_a$                 & DNA alignment length & $15000$ \\
    L                     & MCMC chain length & 1M \\
    $R_i$                 & RNG seed incipient tree & 1 to 40000 \\
    $R_a$                 & RNG seed alignment simulation & 1 to 40000 \\
    $R_b$                 & RNG seed BEAST2 & 1 to 40000 \\
    \hline
  \end{tabular}
  \caption{
    Overview of the 12 simulation parameters.
  }
  \label{table:simulation_parameters}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{l l}
    \hline
    $n$ & Description \\
    \hline
    \hline
    $12$   & simulation parameters \\
    $1000$ & nLTT statistic \\
    \hline
    \hline
    $12$   & simulation parameters \\
    $1$ & [NOTE: add yes/no?] expected speciation completion time (from Etienne \& Rosindell 2012) \\
    $1$ & [NOTE: add yes/no?] expected duration of speciation (from Etienne \& Rosindell 2012) \\
    $1$ & number of taxa in incipient species tree (also: number of subspecies) \\
    $1$ & [NOTE: add yes/no?] difference in branch lengths between the two sampling methods? \\
    $1$ & number of taxa in species tree (also: number of species) \\
    $1$ & mean nLTT statistic \\
    $1$ & median nLTT statistic \\
    $1$ & lower bound of 95\% HPD value of nLTT statistic \\
    $1$ & upper bound of 95\% HPD value of nLTT statistic \\
    $11$ & ESSes of all parameters estimated by BEAST2 (see specs below) \\
    $?10$ & MCMC operator acceptances\\
    \hline
  \end{tabular}
  \caption{
    Specification of the data sets. $n$ denotes the number
    of columns a certain item will occupy. Above the double horizontal line,
    the content of the raw data file is specified, resulting in a table of 
    1012 columns and 40K rows. Below the double horizontal line,
    the content of the summary data file specified, resulting in a table of
    42 columns and 40K rows.
  }
  \label{table:raw_data_set_specs}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{l l}
    \hline
    \# & Description \\
    \hline
    \hline
    1 & posterior \\
    2 & likelihood \\
    3 & prior \\
    4 & treeLikelihood \\
    5 & TreeHeight \\
    6 & BirthDeath \\
    7 & BDBirthRate \\
    8 & BDDeathRate \\
    9 & logP.mrca \\
    10 & mrcatime \\
    11 & clockRate \\
    \hline
  \end{tabular}
  \caption{
    Overview of the 11 BEAST2 estimated parameters
  }
  \label{table:estimated_parameters}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
